---
title: "Week 12 - Teacher"
author: "Teacher"
date: "Assignment Due Date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1

1. Install the tidycensus package with `install.packages("tidycensus")`. Be sure to remove this from your R Markdown file; otherwise, it may not knit.
2. Load the package with `library()` in the chunk below.
3. Obtain an API key: <https://api.census.gov/data/key_signup.html>
4. Run `census_api_key("YOUR KEY GOES HERE", install = TRUE)` and replace "YOUR KEY GOES HERE" with your API key. Delete this line after it has successfully run and do **not** leave it in your code.
5. Restart your R session (shortcut: Shift + Command + 0).

Please do **not** leave your API key in your code. If you include your API key in your knitted file, it will result in a 5 point deduction.

```{r}

```

## Step 2




## Rubric

|                                               | Points Possible | Points Earned |
|-----------------------------------------------|-----------------|---------------|
| File knits                                    | 2               |               |
| Code is accurately filled in for all 5 steps. | 10              |               |
| API key is removed from R Markdown file       | 5               |               |
| Step 4 Questions                              | 8               |               |
| Step 5 Questions                              | 7               |               |

## Knit and Submit

Complete the following steps to knit and publish your work (note - we'll remove these steps after this week, as we know you're becoming quite familiar with it!):

1.  First, change the name of the `author:` in the [YAML header](https://bookdown.org/yihui/rmarkdown-cookbook/rmarkdown-anatomy.html#yaml-metadata) at the very top of this document to your name. The YAML header controls the style and feel for knitted document but doesn't actually display in the final output.

2.  Next, click the knit button in the toolbar above to "knit" your R Markdown document to a [HTML](https://bookdown.org/yihui/rmarkdown/html-document.html) file that will be saved in your R Project folder. You should see a formatted webpage appear in your Viewer tab in the lower right pan or in a new browser window. Let's us know if you run into any issues with knitting.

3.  Upload the HTML file that is now present in the same folder as this R Markdown file to Canvas to receive credit for this assignment!
