---
title: "Week 14 Slides"
subtitle: "Topic Modeling"
author: "**Isabella Velásquez and Maryrose Weatherton**"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: xaringan::moon_reader
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r, echo=FALSE}
# then load all the relevant packages
pacman::p_load(pacman, knitr, tidyverse, readxl)
```

```{r xaringanExtra-clipboard, echo=FALSE}
# these allow any code snippets to be copied to the clipboard so they 
# can be pasted easily
htmltools::tagList(
    xaringanExtra::use_clipboard(
        button_text = "<i class=\"fa fa-clipboard\"></i>",
        success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    ),
    rmarkdown::html_dependency_font_awesome()
)
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```


# Purpose and Agenda

This week, we'll dive into an element of data analysis that is surprisingly commonly-needed --- and useful: pivoting. Pivoting describes transforming data from a form in which variables are represented in rows to one in which they are represented in columns. Sometimes, these different forms are described as "long" and "wide."

## What we'll do in this presentation

- Discussion
- Key Concept: tidy data
- Key Concept: Pivoting
- Code-along and data
- Discussion
- What's next: Assignment(s) and Readings

---

# Discussion

.panelset[

.panel[.panel-name[Background]

- There are some conventions that are stated and sometimes unstated about how data should be recorded
- For example, typically, rows represent _cases_ (AKA _observations_), while columns represent _variables_

]

.panel[.panel-name[Discussion Question]

- Which conventions (stated or unstated!) for how data should be recorded have you found to be important in your work? Why?

]

]


---

# Optional Reading Recap

Optional Substantive reading(s):

> Lucy, L., Demszky, D., Bromley, P., & Jurafsky, D. (2020). Content analysis of textbooks via natural language processing: Findings on gender, race, and ethnicity in Texas US history textbooks. AERA Open, 6(3), https://journals.sagepub.com/doi/full/10.1177/2332858420940312

Technical Reading(s):

> (on text analysis)

]

---

# Key Concept: tidy data

.panelset[

.panel[.panel-name[What is tidy data?]

- *Tidy data* refers to a particular expression of the conventions about how data should be recorded, or stored
- The idea was first shared in a *Journal of Statistical Software* paper (that you can optionally read this week!)
- Broadly, "The principles of tidy data provide a standard way to organise data values within a dataset" (Wickham, 2014, p. 1)
- The idea of tidy data relates to other notions how how data should be recorded, including schema for relational databases
]


.panel[.panel-name[Definition and Benefits]

**Definition**:

1. Each variable forms a column.
2. Each observation forms a row.
3. Each type of observational unit forms a table.

**Benefits**:

- Using the "tidy data" gives you a sense of direction for how you can prepare disorganized data
- The "tidy data" format works well with many tidyverse functions, including the `group_by()` and `summarize()` we used last week, as well as ggplot2() functions

]

.panel[.panel-name[Example]

**How can we make this tidy?**

```{r, echo = FALSE, eval = TRUE}
tibble(name = c("Maryrose", "Isabella"),
       observation_at_t1 = c(3, 2),
       observation_at_t2 = c(5, 1),
       observation_at_t3 = c(3, 5))
```

]

.panel[.panel-name[Example]

**Each observation must become a row, with a variable indicating when the observation was made**

```{r, echo = FALSE, eval = TRUE}
tibble(name = rep(c("Maryrose", "Isabella"), 3),
       time = c(1,1,2,2,3,3),
       observation = c(3, 2, 5, 1, 3, 5))
```

*For examples of other common problems with datasets, check out [the tutorial reading for this week](https://vita.had.co.nz/papers/tidy-data.pdf) on tidy data.*
]

]

---

# Key Concept: Pivoting

.panelset[

.panel[.panel-name[How to tidy]

- In the section on the last key concept, we showed a dataset in a tidy form (and its original form)
- We manually changed the format of that dataset, because it was very small in size
- But, how can we tidy data when we're working with larger datasets---when manual transformations are not possible?
- One set of functions to do this* are the pivot functions:
- `pivot_longer()`
- `pivot_wider()`

*There are aspects of tidy data that cannot be addressed through the pivot functions, though we focus on them because the problem they address is very common and otherwise difficult to resolve 
]

.panel[.panel-name[`pivot_longer()`]

- When should `pivot_longer()` be used?
- For the following data frame, this function can be used to tidy the data (we'll show how to use this function in just a moment!)

```{r, echo = FALSE, eval = TRUE}
tibble(name = c("Maryrose", "Isabella"),
       observation_at_t1 = c(3, 2),
       observation_at_t2 = c(5, 1),
       observation_at_t3 = c(3, 5))
```

]

.panel[.panel-name[`pivot_wider()`]

- When should `pivot_wider()` be used?
- For the following data frame, this function can be used to change the data back to its original form

```{r, echo = FALSE, eval = TRUE}
tibble(name = c("Maryrose", "Isabella"),
       observation_at_t1 = c(3, 2),
       observation_at_t2 = c(5, 1),
       observation_at_t3 = c(3, 5))
```

]

.panel[.panel-name[Notes]

- You may be wondering, why would we _ever_ want to use `pivot_wider()`, if doing so typically changes data (back) into non-tidy form?
- Sometimes, non-tidy data is useful:
- For displays of (tables with) data, for which non-tidy data is more familiar or intuitive
- For certain functions or data analytic techniques that require the data to be in this form
]


]

---

# Code-along

.panelset[

.panel[.panel-name[pivot_longer()]

- Let's start by tidying the example data frame
- You can click the copy symbol in the top-right of the below code chunk to easily copy this code
- `pivot_longer()` requires that you specify the `names_to` and `values_to` and we use `!` to indicate which columns should _not_ be pivoted

```{r, echo = TRUE, eval = FALSE}
dat <- tibble(name = c("Maryrose", "Isabella"),
              observation_at_t1 = c(3, 2),
              observation_at_t2 = c(5, 1),
              observation_at_t3 = c(3, 5))

dat %>% 
    pivot_longer(!name, names_to = "time", values_to = "values")
```

]

.panel[.panel-name[pivot_wider()]

- Let's start with our data in long form

```{r, echo = TRUE, eval = FALSE}
dat_long <- dat %>% 
    pivot_longer(!name, names_to = "time", values_to = "values")
```

- How might we change this _back_ into wide form?
- `pivot_wider()` requires you to specify `names_from` and `values_from`

```{r, echo = TRUE, eval = FALSE}
dat_long %>% 
    pivot_wider(names_from = "time", values_from = "values")
```
]

.panel[.panel-name[More practice]

- Let's try tidying a different data frame, transforming it from _wide_ to _long_ form 
- How might we do this? Consider a) which variable(s) should not be pivoted and what the b) `names_to` and c) `values_to` should be

```{r, echo = TRUE, eval = FALSE}
relig_income
```

]
]

---

# What's next?

.panelset[

.panel[.panel-name[Weekly Assignment]

- In this week's assignment, you will tidy a data set while also practicing selecting and renaming variables and joining a second data set

]

.panel[.panel-name[Mini-Project]

- The mini-project is now available in Canvas! You will be using recent data from the Tennessee Department of Education. While the data is specified, the analyses you conduct to answer two questions of the data are more open-ended in nature. You'll have two week to complete this, to allow for time for you to ask questions and dig deep into the data. One pointer - note that many of the variables are at the school-level and represent the percentage of teachers who chose a particular response. How you choose to "wrangle" or prepare the data given the nature of these variables is an important consideration as you work on your mini project. Our goal with this assignment is not perfection, but rather for you to practice capabilities you have developed over the semester. This is reflected in the grading rubric for the assignment!

]

.panel[.panel-name[Data]

- The data we'll use comes from the [*National Teacher and Principal Survey*](https://nces.ed.gov/surveys/ntps/) and is on the number of full-time nurses in schools in U.S. states.

.panel[.panel-name[Optional Substantive Reading(s)]

Substantive reading(s):

> Lemay, D. J., Baek, C., & Doleck, T. (2021). Comparison of learning analytics and educational data mining: A topic modeling approach. Computers and Education: Artificial Intelligence, 2, https://www.sciencedirect.com/science/article/pii/S2666920X21000102

This is linked in Canvas.

]

.panel[.panel-name[Technical Reading(s)]

- Technical reading(s):

> (on topic modeling)

This is also linked in Canvas.

]

]