---
title: "Week 5 Slides"
subtitle: "How do I access and describe data? Using an R package and an API"
author: "**Isabella Velásquez and Maryrose Weatherton**"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)

style_duo_accent(primary_color = "#FF8200",
                 secondary_color = "#58595B",
                 title_slide_text_color = "#222943",
                 title_slide_background_color = "#ededed",
                 title_slide_background_image = "https://brand.utk.edu/wp-content/uploads/2019/02/University-HorizRightLogo-RGB.png",
                 title_slide_background_position = "bottom",
                 title_slide_background_size = "30%")
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r, echo=FALSE}
# then load all the relevant packages
pacman::p_load(pacman, knitr, tidyverse, readxl)
```

```{r xaringanExtra-clipboard, echo=FALSE}
# these allow any code snippets to be copied to the clipboard so they 
# can be pasted easily
htmltools::tagList(
    xaringanExtra::use_clipboard(
        button_text = "<i class=\"fa fa-clipboard\"></i>",
        success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    ),
    rmarkdown::html_dependency_font_awesome()
)
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```


# Purpose and Agenda

Data scientists often talk about "digital data" and "log-trace" data. How do I access such data? This week, we explore ways to do that through Application Programming Interfaces (APIs)---specifically, through the educationdata R package.

## What we'll do in this presentation

- Discussion
- Reading Recap
- Key Concept: Log-trace data and Application Programming Interfaces (APIs)
- Key Concept: The educationdata package
- Code-along and data
- What's next: Assignment(s) and Readings

---

# Discussion

.panelset[

.panel[.panel-name[Background]



]

.panel[.panel-name[Discussion Question]


]

]

---

# Reading Recap

Substantive reading(s):

> Reich, J. (2022). Learning analytics and learning at scale. Lang, C., Siemens, G. Wise, AF, Gašević, D., Merceron, A.,(eds.) Handbook of Learning Analytics, 2nd ed. Vancouver, Canada: SOLAR.

> White, R. S. (2023). Ceilings Made of Glass and Leaving En Masse? Examining Superintendent Gender Gaps and Turnover Over Time Across the United States. Educational Researcher, 0013189X231163139.

Technical reading(s):

> Introducing educationdata, https://urbaninstitute.github.io/education-data-package-r/articles/introducing-educationdata.html (also, check out other pages on this website)

???

As in the first week, it may be good to pull out and summarize highlights from students' discussion posts here.

---

# Course Logistics

We're now two weeks into class! What has gone well? What issues have you had? What questions do you have?

Syllabus: https://rpubs.com/jmichaelrosenberg/stem-680-fall-2023-syllabus

And the weekly schedule: https://docs.google.com/document/d/1cUrX_bpHFNdqvSZU_cw0mwEWuz_3m2Sn7TjIVILAv_E/edit

???

May be a chance to return to the syllabus and schedule and answer any questions that have arisen.

---

.title[Key Concept: ggplot2]

.panelset[

.panel[.panel-name[xxx]


]

.panel[.panel-name[yyyy]


]

.panel[.panel-name[zzz]

]
]

---

# Key Concept: 



---

# Code-along

*Let's code-along together in an `.R` we'll use---just like we did last week*

.panelset[

.panel[.panel-name[ggplot()]

- we use `ggplot()` to start our plotting of the built-in `penguins` dataset from the palmerpenguins package

```{r, eval = FALSE, echo = TRUE}
library(tidyverse) # let's load the tidyverse, loading ggplot2 & other packages
library(palmerpenguins) # make sure to install with install.packages() first

ggplot(penguins)
```

*This won't do... much (yet!)*

]

.panel[.panel-name[aes()]

- add on to our initial code by specifying the _aesthetic mapping_ with `aes()`
- specfically, let's tell R what the *x* variable is - let's pick _any_ numeric variable from the data set

```{r, eval = FALSE, echo = TRUE}
glimpse(penguins) # to see which variables are numeric

ggplot(penguins, aes(x = body_mass_g) # can change to another!
```

**Still not doing much!**

]

.panel[.panel-name[geom_histogram]

- Last, specify the geometric object you want to use to represent the variable you mapped to an aesthetic
- Let's try a histogram!
- We always add *layers* onto ggplot2 plots with the `+` operator

```{r, eval = FALSE, echo = TRUE}
ggplot(penguins, aes(x = body_mass_g) +
           geom_histogram()
```

*Voilà!*

]

.panel[.panel-name[geom_point]

- Let's expand on what we've done by examining _two_ variables, one we "map" to *x*, another to *y*
- We'll see how the bill length and depth of penguins relate, using a different geom to do so with a scatterplot

```{r, eval = FALSE, echo = TRUE}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm) +
           geom_point()
```

*Voilà (take two)!*

]

.panel[.panel-name[A 3rd variable]

- Part of the joy of ggplot2 is the flexibility you have in how you can modify a plot
- What if we want to visualize a third variable? 
    - We have several aesthetic features we can "map" this third variable to, including the size, shape, and color of points

```{r, eval = FALSE, echo = TRUE}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species) +
           geom_point()
```

*Voilà (take three)!*

]

.panel[.panel-name[More additions]

- The sky is the limit! 
- We can also add other geoms to our plot, like one representing the _slope_ of the relation between *x* and *y*
- Furthermore, we can easily _theme_ out plot

```{r, eval = FALSE, echo = TRUE}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
           geom_point() +
           geom_smooth(method = "lm") +
    theme_bw()
```

]

]

---

# What's next?

.panelset[


.panel[.panel-name[Weekly Assignment]

- We'll be doing more with ggplot2 in this lab, while returning back to some of the functions we used last week
- We'll be faceting our plots using a _layer_ we can add, `facet_wrap()`

]

.panel[.panel-name[Data]

- We'll be using a large, complex dataset from the United States Department of Education
- Specifically, we'll use data from the [Elementary/Secondary Information System (ElSi)](https://nces.ed.gov/ccd/elsi/), a data collection system that includes information on the more than 100,000 schools and more than 8,000 school districts in the country
- Note: Even though this data is alrge and complex, it is _not_ cleaned, and one slightly unusual thing we'll need to do is _skip_ the first six rows of the CSV file accessed through the ElSi, because they're blank!

]

.panel[.panel-name[Substantive Reading(s)]

- Substantive reading(s) for the week ahead:

> Shelton, C. C., Koehler, M. J., Greenhalgh, S. P., & Carpenter, J. P. (2022). Lifting the veil on Teach ersP ayTeachers. com: An in vestigation of educational marketplace offerings and downloads. Learning, Media and Technology, 47(2), 268-287.

> Rosenberg, J. M., Borchers, C., Burchfield, M. A., Anderson, D., Stegenga, S. M., & Fischer, C. (2022). Posts about students on Facebook: a data ethics p erspective. Educational researcher, 51(8), 547-550.

Linked in Canvas.

]

.panel[.panel-name[Technical Reading(s)]

- Technical reading(s) for the week ahead:

> R for Data Science, Data Transformation: https://r4ds.hadley.nz/data-transform

Also linked in Canvas.

]

]