---
title: "Week 5 - Teacher"
author: "Teacher"
date: "Assignment Due Date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1

Load the tidyverse and educationdata packages. You may need to install the educationdata package before loading. 

```{r}
library(tidyverse)
library(educationdata)
```

## Step 2

Query the educationdata API to create a data frame that includes fall enrollment for colleges and universities broken down by race and sex for the years 1990, 2000, 2010, and 2020 in the state of Tennessee.

```{r}
#Before I can get this to run, I need the 'ncessch' value (or whatever number is used for colleges because it seems ncessch is for K12). I looked online but I can't figure out how to find it. In the R documentation on GitHub, they have the ncessch value for New Jersey but don't document where they got that from. On the website https://educationdata.urban.org/data-explorer/explorer you can also explore just UTK data. If you know how we can filter 

enroll_race <- get_education_data(level = 'college-university',
                                  source = 'ipeds',
                                  topic = 'fall-enrollment',
                                  subtopic = list('race', 'sex'),
                                  filters = list(year = c(1990, 2000, 2010, 2020),
                                                 ncessch = ___),
                                  add_labels = TRUE)
```

## Step 3

Create a faceted plot comparing enrollemnt over the four selected years. 

```{r}

```

What is one interesting piece of information you learned about college enrollment in Tennessee from your exploration?

- 

## Step 4

Create your own data frame! Examine the options available on the educationdata GitHub (https://github.com/UrbanInstitute/education-data-package-r). 

```{r}

```

Why did you select the variables you selected?

- 

What analysis (graphing, summary statistics, etc.) could you perform to glean meaningful information?

- 

## Step 5

Complete your analysis described in the previous step! If your analysis is too complicated, you can perform a scaled version. 

```{r}

```

What information did you learn from your analysis? Were there any results that surprised you?

- 

- 


## Reach

This portion of the assignment is optional but offers you a chance to push your skills further.

Think of an issue Donde Plowman might be interested in. Create a new data frame that could be presented to Donde Plowman to address this interest. Complete a simple analysis to hep Donde understand what the data is saying. 

```{r}

```

## Rubric

|                                               | Points Possible | Points Earned |
|-----------------------------------------------|-----------------|---------------|
| File knits                                    | 2               |               |
| Code is accurately filled in for all 5 steps. | 10              |               |
| Step 3 Question                               | 3               |               |
| Step 4 Questions                              | 8               |               |
| Step 5 Question                               | 7               |               |

## Knit and Submit

Complete the following steps to knit and publish your work:

1.  First, change the name of the `author:` in the [YAML header](https://bookdown.org/yihui/rmarkdown-cookbook/rmarkdown-anatomy.html#yaml-metadata) at the very top of this document to your name. The YAML header controls the style and feel for knitted document but doesn't actually display in the final output.

2.  Next, click the knit button in the toolbar above to "knit" your R Markdown document to a [HTML](https://bookdown.org/yihui/rmarkdown/html-document.html) file that will be saved in your R Project folder. You should see a formatted webpage appear in your Viewer tab in the lower right pan or in a new browser window. Let's us know if you run into any issues with knitting.

3.  Upload the HTML file that is now present in the same folder as this R Markdown file to Canvas to receive credit for this assignment!
