---
title: "Week 9 Assignment"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1

Here, please load the tidyverse and educationdata packages.

```{r}
library(tidyverse)
library(educationdata)
```

## Step 2

```{r}

```



|                                               | Points Possible | Points Earned |
|-----------------------------------------------|-----------------|---------------|
| The three reflective steps (2, 3, 8) complete | 7.5 each        |               |
| Coding steps (1, 4, 5, 6, 7) complete         | 7.5 each.       |               |

## Knit and Submit

Complete the following steps to knit and publish your work:

1.  First, change the name of the `author:` in the [YAML header](https://bookdown.org/yihui/rmarkdown-cookbook/rmarkdown-anatomy.html#yaml-metadata) at the very top of this document to your name. The YAML header controls the style and feel for knitted document but doesn't actually display in the final output.

2.  Next, click the knit button in the toolbar above to "knit" your R Markdown document to a [HTML](https://bookdown.org/yihui/rmarkdown/html-document.html) file that will be saved in your R Project folder. You should see a formatted webpage appear in your Viewer tab in the lower right pan or in a new browser window. Let's us know if you run into any issues with knitting.

3.  Upload the HTML file that is now present in the same folder as this R Markdown file to Canvas to receive credit for this assignment!
