---
title: "Week 9 Slides"
subtitle: "Group By and Summarize"
author: "**Isabella Velásquez and Maryrose Weatherton**"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)

style_duo_accent(
    primary_color = "#FF8200",
    secondary_color = "#58595B",
    title_slide_text_color = "#222943",
    title_slide_background_color = "#ededed",
    title_slide_background_image = "https://brand.utk.edu/wp-content/uploads/2019/02/University-HorizRightLogo-RGB.png",
    title_slide_background_position = "bottom",
    title_slide_background_size = "30%"
)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(palmerpenguins)
```

```{r, echo=FALSE}
# then load all the relevant packages
pacman::p_load(pacman, knitr, tidyverse, readxl)
```

```{r xaringanExtra-clipboard, echo=FALSE}
# these allow any code snippets to be copied to the clipboard so they 
# can be pasted easily
htmltools::tagList(
    xaringanExtra::use_clipboard(button_text = "<i class=\"fa fa-clipboard\"></i>",
                                 success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>", ),
    rmarkdown::html_dependency_font_awesome()
)
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

# Purpose and Agenda

What do you need to be able to work with a range of datasets? This week focuses on core data wrangling skills; it is the first of four weeks that equip you with a set of skills focused on the dplyr and tidyr packages that are part of the tidyverse. The context is accessing and preparing data on colleges in advance of building a better rating system.

## What we'll do in this presentation

- Reading Recap
- Discussion
- Key Concept: Group by and summarize
- Code-along and data
- What's next: Assignment(s) and Readings

---

# Reading Recap

Substantive reading(s):

> Morse, R., and & Brooks, E. (2023). How U.S. News Calculated the 2024 Best Colleges Rankings. https://www.usnews.com/education/best-colleges/articles/how-us-news-calculated-the-rankings

> Wolfe, R. (2023). A different kind of college ranking. *Washington Monthly.* https://washingtonmonthly.com/2023/08/27/a-different-kind-of-college-ranking-2/

> Blinder, A. (2023). With a New Formula, U.S. News  Rankings Boost Some State Universities. *New York Times*. https://www.nytimes.com/2023/09/18/us/us-news-college-ranking.html

Technical reading(s):

> Esposito, E., & Stark, D. (2019). What's Observed in a Rating? Rankings as Orientation in the Face of Uncertainty. *Theory, Culture & Society, 36*(4), 3-26.

---

# Discussion

.panelset[

.panel[.panel-name[Background]

- One way we can think about _large and complex_ data sets is whether the data is aggregated---or not
- In education data, it is highly common to aggregate raw data (e.g., individual student assessments) at multiple levels:

- Multiple assessments at the **student level**
- Multiple students at the **teacher level**
- Multiple teachers at the **subject level** and **school level**
- Multiple schools at the **district level**
- Multiple districts at the **state level**

- What counts as raw and aggregated data differs for each field---and even for each research question!

]

.panel[.panel-name[Discussion Questions]

- For your research-related area of interest, what are one or more levels at which raw data is aggregated?
- For which purposes might **raw data** me more useful? For which purposes might aggregate data be useful?

]

]

---

# Key Concept: Aggregating


.panelset[

.panel[.panel-name[dplyr "verbs"]

- As we discussed in our last class (for weeks 7-8), there are several functions loaded in the dplyr package (which itself is loaded in the tidyverse package) that are useful in practically all analyses: `select()`, `rename()`, `filter()`, `arrange()`, and `mutate()`
- Several additional "verbs" are also very useful and are the focus of this and the next week: `group_by()`, `summarize()`, `pivot_longer()`, and `pivot_wider()`
- We focus on the first two of these additional "verbs" this week
- We use the term "aggregating" to refer to grouping and summarizing - specifying groups in our data and summarizing them
]

.panel[.panel-name[group_by]

- `group_by()` is used to specify the group or groups in a a data frame (or tibble)
- It works by specifying the variables you want to keep or remove
- Let's imagine we're using school-level data on the number of students enrolled

```{r, echo = TRUE, eval = FALSE}
grouped_data <- dat %>%
    group_by(district_name)

grouped_data # the data is now "grouped" by district_name
```

- *Important reminder from last week!*: if we don't assign the output a new name (`selected_data <-`), the results of our use of a dplyr "verb" won't be saved

]

.panel[.panel-name[multiple groups]

- Importantly, we can specify multiple groups; typically, groups are a character string or factor, but they can be any unique identifiers

```{r, echo = TRUE, eval = FALSE}
grouped_data_2 <- dat %>%
    group_by(district_name, state_name)

grouped_data_2 # the data is now "grouped" by school *and* district both
```

What this grouping means is that **summary statistics will be calculated for each unique combination of groups** (specific schools within specific districts and states, even if the schools share the same name)

]

.panel[.panel-name[summarize]

- Once we have specified groups, we can then calculate summary statistics
- The first part of the `summarize()` function is simply the name of our summary statistic --- whatever we want it to be!
- The second part, following the `=` symbol, is the formula for how your summary statistic is calculated (as long as it returns a single value per group, it is a valid summary statistic!)
- This sounds pretty abstract, but it can also be pretty straightforward, as in the following:

```{r, echo = TRUE, eval = FALSE}
grouped_data %>% 
    summarize(mean_number_of_students_enrolled = mean(number_of_students_enrolled)) 
```
]

.panel[.panel-name[piping it!]

- We can use the pipe operator to link together these functions in a single block of code
- We can also change our summary statistic or add multiple summary statistics, as below
- Practically always, summary statistics are calculated for numeric variables

```{r, echo = TRUE, eval = FALSE}
dat %>% 
    group_by(district_name, state_name) %>% 
    summarize(mean_number_of_students_enrolled = mean(number_of_students_enrolled),
              max_number_of_students_enrolled = max(number_of_students_enrolled),
              range_number_of_students_enrolled = range(number_of_students_enrolled)) 
```

]
]

---

# Code-along

.panelset[

.panel[.panel-name[grouping]

```{r, eval = FALSE, echo = TRUE}
library(palmerpenguins)
library(tidyverse)

grouped_data <- penguins %>%
    group_by(species)

grouped_data # the data is now "grouped" by species
```

]

.panel[.panel-name[Multiple groups]

```{r, eval = FALSE, echo = TRUE}
grouped_data_2 <- penguins %>%
    group_by(species, island)

grouped_data # the data is now "grouped" by species and island
```

]

.panel[.panel-name[summarize]

```{r, eval = FALSE, echo = TRUE}
grouped_data %>% 
    summarize(mean_body_weight_g = mean(body_weight_g))

# note: often, we need to add na.rm = TRUE to remove rows with missing values
# can you replace mean with a different measure of central tendency? what do you find?

grouped_data_2 %>% 
    summarize(mean_body_weight_g = mean(body_weight_g)) # what is different in this output?

grouped_data_2 %>% 
    summarize(mean_body_weight_g = mean(body_weight_g),
              max_body_weight_g = max(body_weight_g),
              n = n()) # n() is a helpful summary function that returns the number of rows per group

```

]

.panel[.panel-name[Doing more]

- There are some helpful extensions of summarize that may prove useful
- Imagine you want to find the mean for _all_ of the numeric variables in your data, after you have gruped by species and island
- Let's start from the raw data, building this up step-by-step, and assigning the output a new name

```{r, eval = FALSE, echo = TRUE}
mean_vals_for_penguins <- penguins %>% 
    group_by(species, island) %>% 
    summarize_if(is.numeric, mean, na.rm = TRUE) # what if we don't add na.rm = TRUE? try this out!

mean_vals_for_penguins
```

- What if we wanted to remove year? We can use select to remove it --- or, try out `summarize_at()`:

```{r}
mean_vals_for_penguins <- penguins %>% 
    group_by(species, island) %>% 
    summarize_at(vars(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g), mean, na.rm = TRUE)


mean_vals_for_penguins
```

]

]

---

# What's next?

.panelset[

.panel[.panel-name[Weekly Assignment]

- In the weekly assignment, you will group and summarize school-level data on enrollments, practicing aggregating in three ways:

1. at different levels (groups)
2. for different variables
3. with different summary statistics

]


.panel[.panel-name[Data]

- We'll be using an already-saved version of data we accessed via the educationdata package on school-level enrollments
- The data will be fairly large and complex --- thus suggesting the importance of grouping and summarizing as exploratory data analysis steps

]


.panel[.panel-name[Final Project Planning]

- There is an additional assignment this week: Final Project Planning
- In this assignment, you will read a bit about the final project, and share some _initial, preliminary ideas_ (not final plans!) to receive early feedback (and to get your gears turning regarding what you will do)
- Since this is also due in the next week, the assignment and reading discussion are a slightly more brief than that those you have recently completed; be sure to budget time to start to think about your final project

]


.panel[.panel-name[Substantive Reading(s)]

- Substantive reading(s) for the week ahead:

> Fischer, C., Pardos, Z. A., Baker, R. S., Williams, J. J., Smyth, P., Yu, R., ... & Warschauer, M. (2020). Mining big data in education: Affordances and challenges. *Review of Research in Education, 44*(1), 130-160.

*In this reading, focus on the complexity of data used in education (or your field) and how and why data mining can involve data at different levels or groups**

]

.panel[.panel-name[Technical Reading(s)]

- Technical reading(s) for the week ahead:

> Summarizing Data by Groups: https://r-graphics.org/recipe-dataprep-summarize

> Data Transformation: https://r4ds.hadley.nz/data-transform

For the second reading, please review section 4.5 (from this reading that we already completed)

]

]
