---
title: "Week 7 - Teacher"
author: "Teacher"
date: "Assignment Due Date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1

Load the tidyverse package.

```{r}

```

## Step 2

Read the data frames you created last week.

```{r}

```

## Step 3

Join all data frames from last week into one data frame.

```{r}

```


## Step 4

Now that you have one data frame, we will use the four functions from this week to create a metric to find the best college. As a reminder, the four functions you must use are `mutate()`, `case_when()`, and `if_else()`.

```{r}

```

## Reach

This portion of the assignment is optional but offers you a chance to push your skills further.


```{r}

```

## Rubric

|                                               | Points Possible | Points Earned |
|-----------------------------------------------|-----------------|---------------|
| File knits                                    | 2               |               |
| Code is accurately filled in for all 5 steps. | 10              |               |
| Step 3 Question                               | 3               |               |
| Step 4 Questions                              | 8               |               |
| Step 5 Question                               | 7               |               |

## Knit and Submit

Complete the following steps to knit and publish your work:

1.  First, change the name of the `author:` in the [YAML header](https://bookdown.org/yihui/rmarkdown-cookbook/rmarkdown-anatomy.html#yaml-metadata) at the very top of this document to your name. The YAML header controls the style and feel for knitted document but doesn't actually display in the final output.

2.  Next, click the knit button in the toolbar above to "knit" your R Markdown document to a [HTML](https://bookdown.org/yihui/rmarkdown/html-document.html) file that will be saved in your R Project folder. You should see a formatted webpage appear in your Viewer tab in the lower right pan or in a new browser window. Let's us know if you run into any issues with knitting.

3.  Upload the HTML file that is now present in the same folder as this R Markdown file to Canvas to receive credit for this assignment!
