---
title: "Week 7 - Teacher"
author: "Teacher"
date: "Assignment Due Date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1

Load the tidyverse and educaitondata packages.

```{r}
library(tidyverse)
library(educationdata)
```

## Step 2

US News and Rankings has tasked you with creating a better college ranking system. You are allowed to use any data available from the educationdata package. The focus for this week is to identify 3-4 variables you think would make a better ranking system and create the data frames with said variables. 

What are the 3-4 variables you plan to include? It might be helpful to look at the educationdata GitHub page to see what is available.

-

Now, list the 3-4 *individual* data frames from the educationdata API. 

```{r}

```

## Step 3

Next week, we will be joining these data frames together into one single data frame. Remember, you must have a common "key" across each data frame to join on. That is, there must be a column with the same name for each data frame. We suggest using the institution name as your key. In this step, use `rename()` to create that key. 

```{r}

```

You may not want to include all variables educationdata pulls. In this step, use `select()` to include the variables of interest in each data frame. 

```{r}

```

## Step 3

Join all data frames from last week into one data frame.

```{r}

```


## Step 4

Now that you have one data frame, we will use the four functions from this week to create a metric to find the best college. As a reminder, the four functions you must use are `mutate()`, `case_when()`, and `if_else()`.

```{r}

```

## Reach

This portion of the assignment is optional but offers you a chance to push your skills further.


```{r}

```

## Rubric

|                                               | Points Possible | Points Earned |
|--------------------------------------|-----------------|-----------------|
| File knits                                    | 2               |               |
| Code is accurately filled in for all 4 steps. | 10              |               |
| Step 2 Question                               | 3               |               |
| Step 3 Questions                              | 6               |               |
| Step 4 Code and Question                      | 9               |               |

## Knit and Submit

Complete the following steps to knit and publish your work:

1.  First, change the name of the `author:` in the [YAML header](https://bookdown.org/yihui/rmarkdown-cookbook/rmarkdown-anatomy.html#yaml-metadata) at the very top of this document to your name. The YAML header controls the style and feel for knitted document but doesn't actually display in the final output.

2.  Next, click the knit button in the toolbar above to "knit" your R Markdown document to a [HTML](https://bookdown.org/yihui/rmarkdown/html-document.html) file that will be saved in your R Project folder. You should see a formatted webpage appear in your Viewer tab in the lower right pan or in a new browser window. Let's us know if you run into any issues with knitting.

3.  Upload the HTML file that is now present in the same folder as this R Markdown file to Canvas to receive credit for this assignment!
