---
title: "Week 7 Slides"
subtitle: "Creating a Good College Ranking System"
author: "**Isabella Velásquez and Maryrose Weatherton**"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)

style_duo_accent(primary_color = "#FF8200",
                 secondary_color = "#58595B",
                 title_slide_text_color = "#222943",
                 title_slide_background_color = "#ededed",
                 title_slide_background_image = "https://brand.utk.edu/wp-content/uploads/2019/02/University-HorizRightLogo-RGB.png",
                 title_slide_background_position = "bottom",
                 title_slide_background_size = "30%")
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r, echo=FALSE}
# then load all the relevant packages
pacman::p_load(pacman, knitr, tidyverse, readxl)
```

```{r xaringanExtra-clipboard, echo=FALSE}
# these allow any code snippets to be copied to the clipboard so they 
# can be pasted easily
htmltools::tagList(
    xaringanExtra::use_clipboard(button_text = "<i class=\"fa fa-clipboard\"></i>",
                                 success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",),
    rmarkdown::html_dependency_font_awesome()
)
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```


# Purpose and Agenda

What do you need to be able to work with a range of datasets? This week focuses on core data wrangling skills; it is the first of four weeks that equip you with a set of skills focused on the dplyr and tidyr packages that are part of the tidyverse. The context is accessing and preparing data on colleges in advance of building a better rating system.

## What we'll do in this presentation

- Discussion
- Reading Recap
- Key Concept: select, rename, filter, arrange
- Key Concept: Ratings and rankings
- Code-along and data
- What's next: Assignment(s) and Readings

---

# Discussion

.panelset[

.panel[.panel-name[Background]

- APIs can be very powerful, affording access to large data sets in a convenient manner
- However, they can occasionally be difficult to use to due: a) technical issues, b) challenges around the size or amount of the data being accessed, c) the degree of documentation, among others. 

]

.panel[.panel-name[Discussion Question]

- What issues, if any, did you encounter in the last week using the Education Data Portal API via the educationdata R package? How did you overcome them---or try to overcome them?
- What went well in the last week?

]

]

---

# Reading Recap

Substantive reading(s):

> Rutherford, T., Duck, K., Rosenberg, J. M., & Patt, R. (2022). Leveraging mathematics software data to understand student learning and motivation during the COVID-19 pandemic. *Journal of Research on Technology in Education, 54(sup1)*, S94-S131.

> Rosenberg, J. M., Borchers, C., Burchfield, M. A., Anderson, D., Stegenga, S. M., & Fischer, C. (2022). Posts about students on Facebook: a data ethics perspective. *Educational Researcher, 51*(8), 547-550.

Technical reading(s):

> R for Data Science, Data Transformation: https://r4ds.hadley.nz/data-transform

---

# Key Concept: dplyr "verbs"

.panelset[

.panel[.panel-name[dplyr "verbs"]

- There are several functions loaded in the dplyr package (which itself is loaded in the tidyverse package) that are useful in practically all analyses
- These are all "verbs": words used to describe an action, state, or occurrence ([OxfordLanguage](https://languages.oup.com/google-dictionary-en/))
- The core dplyr "verbs" are select, rename, filter, arrange, and mutate
]

.panel[.panel-name[select]

- Select is used to choose particular variables from a data frame (or tibble)
- It works through your specifying of the variables you want to choose

```{r, eval = FALSE, echo = TRUE}
library(dplyr)
data <- tibble(ID = c(1, 2, 3),
                   Name = c("Alice", "Bob", "Charlie", "Stefanie"),
                   Age = c(25, 30, 22, 39),
                   State = c("TN", "NC", "CA", "TN"))

selected_data <- data %>%
  select(Name, Age)
```

- *Important note and point of emphasis*: if we don't assign the output a new name (`selected_data <-`), the results of our use of a dplyr "verb" won't be saved

]

.panel[.panel-name[rename]

- Rename changes the names of variables

```{r, eval = FALSE, echo = TRUE}
selected_renamed_data <- selected_data %>%
  rename(name = Name, age = Age)
```

- *Note*: You can _select_ and _rename_ in a single step using `select()`:

```{r, eval = FALSE, echo = TRUE}
selected_and_renamed <- data %>% 
    select(name = Name, age = Age) # selects Name and Age and renames them at the same time
```

Because select can be used to rename, rename is rarely used in practice.

Learn more about select and rename [here](https://dplyr.tidyverse.org/reference/select.html), including handy "select helpers"

]

.panel[.panel-name[filter]

- Filter reduces the rows of your data based on _logical conditions_ that you specify

```{r, eval = FALSE, echo = TRUE}
filtered_data <- data %>% 
    filter(Age >= 25)
```

```{r, eval = FALSE, echo = TRUE}
filtered_data <- data %>% 
    filter(State == "TN")
```

Learn more about filter [here](https://dplyr.tidyverse.org/reference/filter.html)

]

.panel[.panel-name[arrange]

- Arrange sorts the rows of your data frame based on numeric or alpha-numeric values

```{r, eval = FALSE, echo = TRUE}
arranged_data <- data %>% 
    arrange(Name)
```

You can reverse the order with `desc()`:

```{r, eval = FALSE, echo = TRUE}
arranged_data <- data %>% 
    arrange(desc(Name))
```

Learn more about arrange [here](https://dplyr.tidyverse.org/reference/arrange.html)

]

.panel[.panel-name[mutate]

- Mutate creates new variables

```{r, eval = FALSE, echo = TRUE}
mutated_data <- data %>% 
    mutate(age_next_year = Age + 1)
```

Learn more about mutate [here](https://dplyr.tidyverse.org/reference/mutate.html)

]

]

---

# Key Concept: Ratings and rankings

.panelset[

.panel[.panel-name[educationdata]

- Reminder: The [educationdata R package](https://urbaninstitute.github.io/education-data-package-r/index.html) facilitates access to the [Education Data Portal's API](https://educationdata.urban.org/documentation/index.html)

```{r, eval = FALSE, echo = TRUE}
df <- get_education_data(level = "schools",
                         source = "ccd",
                         topic = "enrollment",
                         filters = list(year = 2021, grade = 12))
```

*Documentation:*

1. The [R package documentation](https://urbaninstitute.github.io/education-data-package-r/articles/introducing-educationdata.html)
2. The [EDP API documentation](https://educationdata.urban.org/documentation/index.html)
]

.panel[.panel-name[Variables]

- Any data on *colleges*: https://educationdata.urban.org/documentation/colleges.html
- These four sources available via educationdata are likely to be particularly helpful:
    - IPEDS
    - College Scorecard
    - Federal Student Aid
    - National Association of College and University Business Officers

]

.panel[.panel-name[Ranking]

- Creating a ranking involves 1) _accessing data_, 2) _preparing_ different variables, 3) _combining_ them into a single score or rank and 4) _interpreting_ the output
- This will likely involve several dplyr functions:
    - `mutate()`
    - `row_number()`
    - `scale()`
    - `arrange()`

]

.panel[.panel-name[Process]

- Week 7, *this week*: Access and standardize data
- Week 8, *next week (fall break)*: Combine data and present and interpret rankings

]

]

---

# Code-along

.panelset[

.panel[.panel-name[1. Accessing]

- Tuition and fees and fall retention rates
- We'll have to access data on each, standardize them, and combine them to create a ranking

- Let's access the data

```{r, eval = FALSE, echo = TRUE}
library(educationdata)
library(tidyverse)

ratio <- get_education_data(level = "college-university",
    source = "ipeds",
    topic = "student-faculty-ratio",
    filters = list(year = 2020))

retention <- get_education_data(level = "college-university",
    source = "ipeds",
    topic = "fall-retention",
    filters = list(year = 2020))
```

]

.panel[.panel-name[2. Preparing]

```{r, eval = FALSE, echo = TRUE}
ratio <- as_tibble(ratio)

ratio <- ratio %>% 
    select(unitid, student_faculty_ratio) # to simply the data

```

```{r, eval = FALSE, echo = TRUE}
retention <- as_tibble(retention)

retention <- retention %>% 
    filter(ftpt == 99) %>% # let's look at the total of both full (1) and part-time (2) students
    select(unitid, retention_rate) # to simply the data
```

]

.panel[.panel-name[3. Combining]

- As the variables are on different scales, calculating _ranks_ may be helpful -- `min_rank()`, which assigns tied ranks to equivalent values

```{r}
x <- c(5, 1, 3, 2, 2, NA)
min_rank(x)
```

```{r, eval = FALSE, echo = TRUE}
combined_data <- inner_join(ratio, retention, by = c("unitid"))

combined_data <- combined_data %>% 
    mutate(student_faculty_ratio_rank = min_rank(student_faculty_ratio),
           retention_rate_rank = min_rank(retention_rate)) %>% 
    select(unitid, student_faculty_ratio_rank, retention_rate_rank)

final_ranks <- combined_data %>% 
    mutate(total_rank = (student_faculty_ratio_rank + retention_rate_rank) / 2) %>% 
    arrange(total_rank)

final_ranks
```

]

.panel[.panel-name[4. Interpreting]

- We need to know the names of the institutions to be able to present and interpret our results.
- We can access that data, too

```{r, eval = FALSE, echo = TRUE}
directory <- get_education_data(level = "college-university",
    source = "ipeds",
    topic = "directory",
    filters = list(year = 2020))

directory <- directory %>% 
    select(unitid, inst_name)

final_ranks %>% 
    left_join(directory) %>% 
    select(inst_name, total_rank)
```

]

]

---

# What's next?

.panelset[

.panel[.panel-name[Weekly Assignment]

- In the weekly assignment, you will expand on what we did in the code-along in several ways:
    - Researching how _other ranking systems_ work
    - Planning out your approach in terms of _three principles you want to adhere to_ in your analysis
    - Adding at least _three variables to use in your analysis_
    - Suggesting possible improvements in the context of your interpretation of your results
- Big picture, this is challenging: **the goal is to use this as an opportunity to practice all of the skills you have developed through this point in an authentic task**, but it will be challenging, and try to not let the good become the enemy of the perfect 
]

.panel[.panel-name[Data]

- We'll again be using data from the educationdata package --- any of the college-related data: https://educationdata.urban.org/documentation/colleges.html#ipeds_directory
- Again, these four sources available via educationdata are likely to be particularly helpful:
    - IPEDS
    - College Scorecard
    - Federal Student Aid
    - National Association of College and University Business Officers

]

.panel[.panel-name[Substantive Reading(s)]

- Substantive reading(s) for the week ahead (**note: there are three plus an optional reading, but the three required readings are brief!***:

> Morse, R., and & Brooks, E. (2023). How U.S. News Calculated the 2024 Best Colleges Rankings. https://www.usnews.com/education/best-colleges/articles/how-us-news-calculated-the-rankings

> Wolfe, R. (2023). A different kind of college ranking. *Washington Monthly.* https://washingtonmonthly.com/2023/08/27/a-different-kind-of-college-ranking-2/

> Blinder, A. (2023). With a New Formula, U.S. News  Rankings Boost Some State Universities. *New York Times*. https://www.nytimes.com/2023/09/18/us/us-news-college-ranking.html

- Optional reading: 

> Esposito, E., & Stark, D. (2019). What's Observed in a Rating? Rankings as Orientation in the Face of Uncertainty. *Theory, Culture & Society, 36*(4), 3-26.

Linked in Canvas.

]

.panel[.panel-name[Technical Reading(s)]

- Technical reading(s) for the week ahead:

> Data Science in Education Using R, Using School-Level Aggregate Data to Illuminate Educational Inequities: https://datascienceineducation.com/c09

Also linked in Canvas.

]

]
