---
title: "Week 3 Assignment"
author: "Teacher Key"
date: "Assignment due date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.  Load necessary packages

```{r}
library(tidyverse)
library(janitor)
```

2.  Load ELSI data

```{r}
elsi <- read_csv("f23-founds-eds/data/elsi-data.csv", skip = 6)
```

3.  Explore the data and clean column names as necessary

```{r}
glimpse(elsi)
elsi_clean <- clean_names(elsi)
elsi_clean <- elsi_clean %>% 
  mutate(total_students_all_grades_excludes_ae_public_school_2021_22 = as.numeric(total_students_all_grades_excludes_ae_public_school_2021_22))
```

4.  Create a plot comparing multiple groups. Use ggplot2 and `facet_wrap()`.

```{r}
elsi_or <- elsi_clean %>% 
  filter(state_name_public_school_latest_available_year == "Oregon")

elsi_or %>% 
  ggplot(aes(x = total_students_all_grades_excludes_ae_public_school_2021_22)) +
    geom_histogram() +
    facet_wrap(~locale_public_school_2021_22)
```

## Questions

1.  What groups did you compare in step 4? What were you able to determine to be the differences between those two groups by looking at your graphs?

-   

2.  How could a policy maker use the information you gathered (or other information contained in this dataset) to suggest changes? Are graphical representations an effective way to do this?

-   

3.  How might you make your graphs look more "publication ready?" You do not need to include code, just a general idea of what could be improved.

-   

## Rubric

|                                 | Points Possible | Points Earned |
|---------------------------------|-----------------|---------------|
| File knits                      | 2               |               |
| Effective use of `facet_wrap()` | 10              |               |
| Question 1                      | 5               |               |
| Question 2                      | 8               |               |
| Question 3                      | 5               |               |
