---
title: "Week 3 Assignment"
author: "Student"
date: "Assignment Due Date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.  Load necessary packages. You may need to load more than one.

```{r}
library(___)
```

2.  Load ELSI data

```{r}
___ <- read_csv(___)
```

3.  Explore the data and clean column names as necessary

```{r}
___(elsi)

```

4.  Create a plot comparing multiple groups. Use ggplot2 and `facet_wrap()`. We've provided a template for a histogram. If you feel that a different plot type will represent the data you are interested in better, please use that geom type. 

```{r}
___ %>% 
  ggplot(aes(x = ___)) +
    geom_histogram() +
    facet_wrap(~___)
```

## Questions

1.  What groups did you compare in step 4? What were you able to determine to be the differences between those two groups by looking at your graphs?

-   

2.  How could a policy maker use the information you gathered (or other information contained in this dataset) to suggest changes? Are graphical representations an effective way to do this?

-   

3.  How might you make your graphs look more "publication ready?" You do not need to include code, just a general idea of what could be improved.

-   

## Rubric

|                                 | Points Possible | Points Earned |
|---------------------------------|-----------------|---------------|
| File knits                      | 2               |               |
| Effective use of `facet_wrap()` | 10              |               |
| Question 1                      | 5               |               |
| Question 2                      | 8               |               |
| Question 3                      | 5               |               |
