## Assignment Recap: Why use inline code?

Let's say we're writing a report using the `mtcars` data to report on the Toyota Corolla. In R Markdown, we include R code by inserting code chunks then write our text inside/outside of the chunks.

```{r}
#| message: false
library(tidyverse)

mtcars %>% 
    as_tibble(rownames = "cars") %>% 
    filter(cars == "Toyota Corolla")
```

So in our R Markdown document, we write: 

The Toyota Corolla gets 33.9 mpg, has 4 cyl, 71.1 disp, 65 hp...

But what if our code changes? Say we get a new dataset, `mtcars2` with updated data on the Toyota Corolla to use in our report.

We could:

* Load the new `mtcars2` dataset
* Rerun the calculation
* Based on the new output, edit our sentence to say "The Toyota Corolla gets X mpg, has A cyl, B disp, C hp..."

But this can cause some issues!

* What if we forget to update our sentence after using the new data?
* What if we write the wrong numbers in our sentence?
* What if we are reporting on many variables?

```md
The Toyota Corolla gets...
The Porsche 914-2 gets... 
The Camaro Z28 gets... 
The AMC Javelina gets... 
The Ferrari Dino gets... 
The Maserati Bora gets... 
```

To display the result of a calculation in your R Markdown document, you can use inline code, which is enclosed in backticks (```) and preceded by an "r", making it indistinguishable from the surrounding text.

`r 1 + 1`

However, your output has to be a single value! Otherwise, R won't know what to print. This is where `pull()` comes in! The `pull()` function is used to extract a single column from a data frame or tibble.

```{r}
#| echo: true
mtcars %>%
    as_tibble(rownames = "cars") %>%
    filter(cars == "Toyota Corolla") %>%
    pull(mpg)
```


So instead of "hardcoding" the numbers in the sentence, we could write:

The Toyota Corolla gets `r mtcars %>% as_tibble(rownames = "cars") %>% filter(cars == "Toyota Corolla") %>% pull(mpg)` mpg...

Some tips:

- The inline code must be in the same line (i.e., no breaks)
- If you are running into issues, copy the part of the code after the `r` and put it in a code chunk.
    - Does the code run?
    - Is it a single value?
