---
title: "Week 4 - Teacher"
author: "Teacher"
date: "Assignment Due Date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load necessary packages. You may need to download the tidykids package by running `remotes::install_github("jrosen48/tidykids")`.
```{r}
library(tidyverse)
library(tidykids)
library(janitor)
```

2. Save the tidykids and NAEP data frames to your environment.
```{r}
tidykids <- tidykids
naep <- read_csv("naep-data.csv")
```

3. Familiarize yourself with the two data frames
```{r}
glimpse(tidykids)
glimpse(naep)
```

4. Join the two data frames. You will want to preserve as much data as possible. As a reminder, the column you join by must have the same name in each data frame, so you might need to use `rename()`.
```{r}
naep <- naep %>% 
  rename("state" = "Jurisdiction")
full_df <- full_join(tidykids, naep, "state")

full_df <- clean_names(full_df)
```

5. Analysis question - use ggplot?

```{r}

```


## Questions
1. Does it matter the order you list data frames in the `full_join()` command?
