---
title: "Week 4 - Teacher"
author: "Teacher"
date: "Assignment Due Date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1

Load necessary packages (tidyverse, tidykids, and janitor). Before loading tidykids, you may need to download the package by running `remotes::install_github("jrosen48/tidykids")`.

```{r}
library(tidyverse)
library(tidykids)
library(janitor)
```

## Step 2

Save the tidykids and NAEP data frames to your environment.

```{r}
tidykids <- tidykids
naep <- read_csv("naep-data.csv")
```

## Step 3

Familiarize yourself with the two data frames

```{r}
glimpse(tidykids)
glimpse(naep)
```

What is one interesting piece of information you learned about the two datasets from your exploration?

```         
- 
```

## Step 4

Now we will practice joining the two dataframes together based on the column that refers to the state. As a reminder, the column you join by must have the same name in each data frame, so you will need to use `rename()`.

```{r}
naep <- naep %>% 
  rename("state" = "Jurisdiction")
```

Now that there is a column in each dataframe with a matching name, let's practice joining. First, let's use `right_join()`.

```{r}
joined_right_1 <- right_join(tidykids, naep, "state")

joined_right_2 <- right_join(naep, tidykids, "state")
```

What do you notice about these two new dataframes? Is there anything different between `joined_right_1` and `joined_right_2`?

-   

-   

You can also use `left_join()`.

```{r}
joined_left_1 <- left_join(tidykids, naep, "state")
```

What do you notice are the differences between `right_join()` and `left_join()`? Would there be a time you would need to use one over the other?

-   

-   

Finally, we can join the two dataframes with `full_join()`. Explore what happens when you list the tidykids and NAEP datasets in different orders.

```{r}
full_df <- full_join(tidykids, naep, "state") # Note for instructor: this is the preferred join. 

full_df <- full_join(naep, tidykids, "state")
```

How does `full_join()` compare to the other two joins you practiced? What would be the value in using `full_join()` over the other joins? For these two datasets, is there one order that makes sense over the other to use?

-   

-   

-   

## Step 5

Create a scatterplot to examine the relationship between a variable originally in the tidykids dataset and a variable originally in the NAEP dataset. Use the full_join dataframe for this exercise. You must specify both the x- and y- variables when using `geom_point()`. We encourage you to change the theme and color of the plot. 

```{r}

```

What interesting relationship (if any) did you find?

-   

## Reach

This portion of the assignment is optional but offers you a chance to push your skills further.

- 

## Rubric

|                                               | Points Possible | Points Earned |
|-----------------------------------------------|-----------------|---------------|
| File knits                                    | 2               |               |
| Code is accurately filled in for all 5 steps. | 10              |               |
| Step 3 Question                               | 3               |               |
| Step 4 Questions                              | 8               |               |
| Step 5 Question                               | 7               |               |

## Knit and Submit

Complete the following steps to knit and publish your work:

1.  First, change the name of the `author:` in the [YAML header](https://bookdown.org/yihui/rmarkdown-cookbook/rmarkdown-anatomy.html#yaml-metadata) at the very top of this document to your name. The YAML header controls the style and feel for knitted document but doesn't actually display in the final output.

2.  Next, click the knit button in the toolbar above to "knit" your R Markdown document to a [HTML](https://bookdown.org/yihui/rmarkdown/html-document.html) file that will be saved in your R Project folder. You should see a formatted webpage appear in your Viewer tab in the lower right pan or in a new browser window. Let's us know if you run into any issues with knitting.

3.  Upload the HTML file that is now present in the same folder as this R Markdown file to Canvas to receive credit for this assignment!
