---
title: "Week 13 Assignment"
author: "Student Name"
date: "Assignment Due Date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this homework, we will work with text gathered from a recent class recording. When Zoom records, it collects a transcript with time stamps and speaker identifiers. Before beginning, please download the Zoom data file(s) from Canvas and save them in a separate folder.

## Step 1

Load the tidyverse package. We will also use the tidytext package. Install and load this package if you have not already done so.

```{r}

```

## Step 2

Let's start by working with some text. Here's an example that captures the core steps.

```{r}
my_sentences <- tibble(sentences = c("I think that she's really good at playing soccer, but he's probably a better dancer, especially at his age", "I really like to watch soccer games, especially for OneKnox", "She was a great goalkeeper, perhaps because she played soccer from such a young age"))

my_sentences %>% 
  unnest_tokens(___, ___) %>% # separate words, removing punctuation
  count(___) %>% # counting words
  ____(desc(n)) %>% # arranging by frequency
  ___(___) # removing common words, or "stop" words
```

Before moving on to analyzing class data, what do you think are two benefits of analyzing text data in your field?

-   

## Step 3

Now, let's do the same with our Zoom data.

```{r}
# First, read the class Zoom transcript CSV data. You can download this from Canvas!

# In this next step, use similar code from Step 2.
transcript %>% 
  unnest_tokens(word, ____) %>% 
  count(____) %>% 
  ____(desc(n)) %>% 
  ____(____)
```

Does looking at the top 10 words spoken in class give you an idea of what the class is about? Or any other meaningful information about the interactions that occurred?

-   

## Step 4

Let's create a column chart of the top words, styling it and making it easier to interpret by using color to enhance the plot. Please use a color package other than the default color and include axis titles. You may have to use `slice()` to make the graph clearer. 

```{r}

```

## Step 5

Calculate the mean duration of the time spoken, then create a plot, coloring the points by the number of times each individual spoke. Be sure to include axis titles.

```{r}
mean_utterance_duration_and_n_utterances <- transcript %>%
    mutate(
        start = lubridate::seconds_to_period(hms(start)),
        end = lubridate::seconds(hms(end)),
        utterance_duration = end - start
    ) %>% 
  group_by(name) %>% 
  summarize(mean_utterance_duration = mean(utterance_duration),
            n_utterances = n()) %>% 
  arrange(desc(mean_utterance_duration))
```

```{r}
# Use this code chunk for the plot.

```

Who spoke the most? Does this result surprise you?

-   

If someone not associated with the class were given your graph, what do you think they would interpret about interactions in the class?

-   

## Step 6

In this step, determine how many times the instructor spoke the word "data" over the class. Hint: use code similar to Steps 2 and 3. As a reach, calculate the mean number of "data" words per minute spoken (you may need to use the package lubridate). 

```{r}

```


## Ethics

What could be an ethical concern with presenting the Zoom data as it was given to you? (e.g., the inclusion of screen names, etc.)

-   

What ethical concerns might be present when sharing word frequencies to a large audience?

-   

What is one other ethical concern you have with text/transcript data?

-   

## Reach

As a reach, conduct a sentiment analysis on the Zoom transcript data from class. To see code on how to do this, visit: https://www.tidytextmining.com/sentiment

```{r}

```

## Rubric

|                                               | Points Possible | Points Earned |
|----------------------------------------|----------------|----------------|
| File knits                                    | 2               |               |
| Code is accurately filled in for all 6 steps. | 10              |               |
| Step 2 Question                               | 3               |               |
| Step 3 Questions                              | 3               |               |
| Step 5 Question                               | 3               |               |
| Ethics Questions (3 points each)              | 9               |               |

## Knit and Submit

Complete the following steps to knit and publish your work:

1.  First, change the name of the `author:` in the [YAML header](https://bookdown.org/yihui/rmarkdown-cookbook/rmarkdown-anatomy.html#yaml-metadata) at the very top of this document to your name. The YAML header controls the style and feel for knitted document but doesn't actually display in the final output.

2.  Next, click the knit button in the toolbar above to "knit" your R Markdown document to a [HTML](https://bookdown.org/yihui/rmarkdown/html-document.html) file that will be saved in your R Project folder. You should see a formatted webpage appear in your Viewer tab in the lower right pan or in a new browser window. Let's us know if you run into any issues with knitting.

3.  Upload the HTML file that is now present in the same folder as this R Markdown file to Canvas to receive credit for this assignment!
