---
title: "Week 13 Slides"
subtitle: "Intro to Text Analysis"
author: "**Isabella Velásquez and Maryrose Weatherton**"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: xaringan::moon_reader
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)

state_by_region <- read_csv("https://github.com/cphalpert/census-regions/blob/master/us%20census%20bureau%20regions%20and%20divisions.csv?raw=true") # data on U.S. states' regions

dat <-
    read_csv(
        "https://github.com/utkeds/f23-founds-eds/blob/main/ntps2021_7341_s1s.csv?raw=true",
        skip = 5
    ) %>% 
  select(State,
         ave_nurses = 2,
         students_per_nurse = 4) %>% 
    filter(!is.na(State)) %>% 
    pivot_longer(-State, names_to = "variable", values_to = "values") %>% 
    left_join(state_by_region)
```

```{r, echo=FALSE}
# then load all the relevant packages
pacman::p_load(pacman, knitr, tidyverse, readxl)
```

```{r xaringanExtra-clipboard, echo=FALSE}
# these allow any code snippets to be copied to the clipboard so they 
# can be pasted easily
htmltools::tagList(
    xaringanExtra::use_clipboard(
        button_text = "<i class=\"fa fa-clipboard\"></i>",
        success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    ),
    rmarkdown::html_dependency_font_awesome()
)
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

# Purpose and Agenda

This week, we'll ...

## What we'll do in this presentation

- Assignment ___ review
- Useful functions for analyzing text
- New text analysis package: tidytext

---

## Assignment ___ review

.panelset[

.panel[.panel-name[*add text here*]

* *add text here*


]


]


---

## tidytext Package

.panelset[

.panel[.panel-name[tidytext]

*text here about tidytext package -- explain key vocab like "token"*


]

.panel[.panel-name[`unnest_tokens()`]

**Transforming text into tidy text:** we must break text into individual tokens. That is, each word must have its own line.

- The first argument you must specify is the new column name the text will be unnested into.

- The second argument is the input column the text comes from.

```{r}
text <- c("Because I could not stop for Death -",
          "He kindly stopped for me -",
          "The Carriage held but just Ourselves -",
          "and Immortality")

library(dplyr)
text_df <- tibble(line = 1:4, text = text)
```

```{r}
text_df_unnest <- text_df %>%
  unnest_tokens(word, text)
```

]

.panel[.panel-name[`anti_join()`]

**Removing Stop Words:** `anti_join()` works to remove observations from a specified dataset. Here we will use the `stop_words` dataset from the tidytext package.

- Stop words are words that are not useful for text analysis (e.g., the, a, and, etc.)

```{r}
text_df_unnest %>% 
    anti_join(stop_words)
```

]

.panel[.panel-name[Other analyses]

*discuss other analysis techniques tidytext can use like sentiment analysis, etc.*

]
]

---

## Code-Along

.panelset[
.panel[.panel-name[Intro]

We will practice text analysis on some simple text before you perform text analysis on our class' Zoom transcript data.

- Please download this week's homework for the code-along.

]
]

---

# Notes on the assignment

- You will need to download the .csv transcript file located in this week's assignment module on Canvas.

---

# What's next?

- 

---

# Optional Reading Recap

Substantive reading(s):



Technical reading(s):


