---
title: "Week 6 - Teacher"
author: "Teacher"
date: "Assignment Due Date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1

Load the tidyverse and educationdata packages. Like all packages, you will need to install the educationdata package before loading it for the first time.

```{r}
library(tidyverse)
library(educationdata)
```

## Step 2

Next, we'll query the educationdata API to create a data frame that includes fall enrollment for colleges and universities for years ten years apart.

*Note -- this may take a few minutes (or more) to run! If it stops running, you may need to try again. If you have tried several times and cannot access the data, please message the instructors.*

```{r}
df <- get_education_data(level = "college-university",
    source = "ipeds",
    topic = "enrollment-headcount",
    filters = list(year = c(2018), level_of_study = "undergraduate"))

df <- as_tibble(df)
```

In the next code chunk, please glimpse the data you accessed.

```{r}
df %>% glimpse()
```

You may notice that instead of College or University names, you have an identifier that is a number, `unitid`. You can look up their names here: https://nces.ed.gov/ipeds/datacenter/InstitutionByName.aspx?stepId=1&sid=fb7e4c13-4a9e-43a5-89eb-dfead5e01d78&rtid=6

You can also use that link (above) to lookup the unitid for a school. Please find the University of Tennessee, Knoxville, unit ID. Then, below, use the `filter()` function to filter the data you already accessed to only include rows associated with the University of Tennessee, Knoxville. Here's how that would be done.

```{r}
df %>% filter(unitid == 221759)
```

## Step 3

How many colleges/universities are there in the data set? Please count up the number of unique colleges/universities with code in the next code chunk. 

```{r}
df %>% count(unitid)
```

## Step 4

To visualize the range of enrollments at all U.S. schools, let's create a histogram of total enrollment. To do this, we need to check out the codebook: https://educationdata.urban.org/documentation/colleges.html#ipeds-enrollment-headcount

Note that values of 99 have a special meaning. What does it mean? Add a note below:

- Total

Check out the `headcount` variable in the codebook to understand its meaning. Then, filter the data so that race and sex are both equivalent to 99. If you've successfully filtered the data, you should have only one row per unitid (college or university).

Then, create a histogram of the headcount variable _after_ you have filtered the data. 

Read about geom_histogram() here: https://ggplot2.tidyverse.org/reference/geom_histogram.html

```{r}
df %>% 
    filter(race == 99 & age == 99) %>% 
    ggplot(aes(x = headcount)) +
    geom_histogram()
```

Please interpret the graph you created. What do you notice? What does this tell you?

- 

## Step 5

Create your own data frame *using a different variable of your choosing*. Examine the options available on the educationdata GitHub: https://github.com/UrbanInstitute/education-data-package-r. Also, consider reviewing the EDP documentation: https://educationdata.urban.org/documentation/index.html. 

At this stage, we are focused on becoming familiar with accessing data from the API, so don't worry too much about the specifics of your selection.

```{r}

```

Why did you select the variables you selected?

- 

## Step 6

Complete a very brief, exploratory analysis of your date. This can be _very_ simple; just learn something about the data you accessed. recall again that the purpose of this step is to get familiar with accessing data from the API.

```{r}

```

What information did you learn from your analysis? Were there any results that surprised you?

- 

- 


## Reach

This portion of the assignment is optional but offers you a chance to push your skills further.

Think of an issue University of Tennessee, Knoxville Chancellor Donde Plowman might be interested in. Create a new data frame that could be presented to Donde Plowman to address this interest. Complete a simple analysis to hep Donde understand what the data is saying. 

```{r}

```

## Rubric

|                                               | Points Possible | Points Earned |
|-----------------------------------------------|-----------------|---------------|
| File knits                                    | 2               |               |
| Code is accurately filled in for all 5 steps. | 10              |               |
| Steps 1-6 are completed                       | 3 points each   |               |

## Knit and Submit

Complete the following steps to knit and publish your work:

1.  First, change the name of the `author:` in the [YAML header](https://bookdown.org/yihui/rmarkdown-cookbook/rmarkdown-anatomy.html#yaml-metadata) at the very top of this document to your name. The YAML header controls the style and feel for knitted document but doesn't actually display in the final output.

2.  Next, click the knit button in the toolbar above to "knit" your R Markdown document to a [HTML](https://bookdown.org/yihui/rmarkdown/html-document.html) file that will be saved in your R Project folder. You should see a formatted webpage appear in your Viewer tab in the lower right pan or in a new browser window. Let's us know if you run into any issues with knitting.

3.  Upload the HTML file that is now present in the same folder as this R Markdown file to Canvas to receive credit for this assignment!
