---
title: "Week 6 - Teacher"
author: "Teacher"
date: "Assignment Due Date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1

Load the tidyverse and educationdata packages. Like all packages, you will need to install the educationdata package before loading it for the first time.

```{r}
library(tidyverse)
library(educationdata)
```

## Step 2

Next, we'll query the educationdata API to create a data frame that includes fall enrollment for colleges and universities broken down by race and sex for the years 1990, 2000, 2010, and 2020 in the state of Tennessee.

```{r}
enroll_race <- get_education_data(level = 'college-university',
                                  source = 'ipeds',
                                  topic = 'fall-enrollment',
                                  subtopic = list('race', 'sex'),
                                  filters = list(year = c(1990, 2000, 2010, 2020)),
                                  add_labels = TRUE,
                                  csv = TRUE)
```

## Step 3

Create a faceted plot comparing enrollment over the four selected years. 

```{r}

```

What is one interesting piece of information you learned about college enrollment in Tennessee from your exploration?

- 

## Step 4

Create your own data frame *using a different variable of your choosing*. Examine the options available on the educationdata GitHub: https://github.com/UrbanInstitute/education-data-package-r. Also, consider reviewing the EDP documentation: https://educationdata.urban.org/documentation/index.html. 

At this stage, we are focused on becoming familiar with accessing data from the API, so don't worry too much about the specifics of your selection.

```{r}

```

Why did you select the variables you selected?

- 

What analysis (graphing, summary statistics, etc.) could you perform to glean meaningful information?

- 

## Step 5

Complete your analysis described in the previous step! If your analysis is too complicated, you can perform a scaled-down version; recall again that the purpose of this step is to get familiar with accessing data from the API.

```{r}

```

What information did you learn from your analysis? Were there any results that surprised you?

- 

- 


## Reach

This portion of the assignment is optional but offers you a chance to push your skills further.

Think of an issue Donde Plowman might be interested in. Create a new data frame that could be presented to Donde Plowman to address this interest. Complete a simple analysis to hep Donde understand what the data is saying. 

```{r}

```

## Rubric

|                                               | Points Possible | Points Earned |
|-----------------------------------------------|-----------------|---------------|
| File knits                                    | 2               |               |
| Code is accurately filled in for all 5 steps. | 10              |               |
| Step 3 Question                               | 3               |               |
| Step 4 Questions                              | 8               |               |
| Step 5 Question                               | 7               |               |

## Knit and Submit

Complete the following steps to knit and publish your work:

1.  First, change the name of the `author:` in the [YAML header](https://bookdown.org/yihui/rmarkdown-cookbook/rmarkdown-anatomy.html#yaml-metadata) at the very top of this document to your name. The YAML header controls the style and feel for knitted document but doesn't actually display in the final output.

2.  Next, click the knit button in the toolbar above to "knit" your R Markdown document to a [HTML](https://bookdown.org/yihui/rmarkdown/html-document.html) file that will be saved in your R Project folder. You should see a formatted webpage appear in your Viewer tab in the lower right pan or in a new browser window. Let's us know if you run into any issues with knitting.

3.  Upload the HTML file that is now present in the same folder as this R Markdown file to Canvas to receive credit for this assignment!
