---
title: "Week 6 Slides"
subtitle: "How do I access and describe data? Using an R package and an API"
author: "**Isabella Velásquez and Maryrose Weatherton**"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)

style_duo_accent(primary_color = "#FF8200",
                 secondary_color = "#58595B",
                 title_slide_text_color = "#222943",
                 title_slide_background_color = "#ededed",
                 title_slide_background_image = "https://brand.utk.edu/wp-content/uploads/2019/02/University-HorizRightLogo-RGB.png",
                 title_slide_background_position = "bottom",
                 title_slide_background_size = "30%")
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r, echo=FALSE}
# then load all the relevant packages
pacman::p_load(pacman, knitr, tidyverse, readxl)
```

```{r xaringanExtra-clipboard, echo=FALSE}
# these allow any code snippets to be copied to the clipboard so they 
# can be pasted easily
htmltools::tagList(
    xaringanExtra::use_clipboard(
        button_text = "<i class=\"fa fa-clipboard\"></i>",
        success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    ),
    rmarkdown::html_dependency_font_awesome()
)
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```


# Purpose and Agenda

Data scientists often talk about "digital data" and "log-trace" data. How do I access such data? This week, we explore ways to do that through Application Programming Interfaces (APIs)---specifically, through the educationdata R package.

## What we'll do in this presentation

- Discussion
- Reading Recap
- Key Concept: Log-trace data and Application Programming Interfaces (APIs)
- Key Concept: The educationdata package
- Code-along and data
- What's next: Assignment(s) and Readings

---

# Discussion

.panelset[

.panel[.panel-name[Background]



]

.panel[.panel-name[Discussion Question]


]

]

---

# Reading Recap

Substantive reading(s):

> Reich, J. (2022). Learning analytics and learning at scale. Lang, C., Siemens, G. Wise, AF, Gašević, D., Merceron, A.,(eds.) Handbook of Learning Analytics, 2nd ed. Vancouver, Canada: SOLAR.

> White, R. S. (2023). Ceilings Made of Glass and Leaving En Masse? Examining Superintendent Gender Gaps and Turnover Over Time Across the United States. Educational Researcher, 0013189X231163139.

Technical reading(s):

> Introducing educationdata, https://urbaninstitute.github.io/education-data-package-r/articles/introducing-educationdata.html (also, check out other pages on this website)

???

As in the first week, it may be good to pull out and summarize highlights from students' discussion posts here.

---

.title[Key Concept: log-trace data and APIs]

.panelset[

.panel[.panel-name[xxx]


]

.panel[.panel-name[yyyy]


]

.panel[.panel-name[zzz]

]
]

---

# Key Concept: educationdata]

.panelset[

.panel[.panel-name[xxx]


]

.panel[.panel-name[yyyy]


]

.panel[.panel-name[zzz]

]
]

---

# Code-along

.panelset[

.panel[.panel-name[xxxx]

- yyy
-zzz

```{r, eval = FALSE, echo = TRUE}

```

]

.panel[.panel-name[xxxx]

- yyy
-zzz

```{r, eval = FALSE, echo = TRUE}

```

]

.panel[.panel-name[xxxx]

- yyy
-zzz

```{r, eval = FALSE, echo = TRUE}

```

]

]

---

# What's next?

.panelset[

.panel[.panel-name[Weekly Assignment]

- 
- 

]

.panel[.panel-name[Data]

- We'll be using data from the educationdata package, specifically

]

.panel[.panel-name[Substantive Reading(s)]

- Substantive reading(s) for the week ahead:

> Shelton, C. C., Koehler, M. J., Greenhalgh, S. P., & Carpenter, J. P. (2022). Lifting the veil on Teach ersP ayTeachers. com: An in vestigation of educational marketplace offerings and downloads. Learning, Media and Technology, 47(2), 268-287.

> Rosenberg, J. M., Borchers, C., Burchfield, M. A., Anderson, D., Stegenga, S. M., & Fischer, C. (2022). Posts about students on Facebook: a data ethics p erspective. Educational researcher, 51(8), 547-550.

Linked in Canvas.

]

.panel[.panel-name[Technical Reading(s)]

- Technical reading(s) for the week ahead:

> R for Data Science, Data Transformation: https://r4ds.hadley.nz/data-transform

Also linked in Canvas.

]

]