---
title: "Week 6 Slides"
subtitle: "How do I preparing data for analysis: Foundational skills, Part 1"
author: "**Isabella Velásquez and Maryrose Weatherton**"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)

style_duo_accent(primary_color = "#FF8200",
                 secondary_color = "#58595B",
                 title_slide_text_color = "#222943",
                 title_slide_background_color = "#ededed",
                 title_slide_background_image = "https://brand.utk.edu/wp-content/uploads/2019/02/University-HorizRightLogo-RGB.png",
                 title_slide_background_position = "bottom",
                 title_slide_background_size = "30%")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r, echo=FALSE}
# then load all the relevant packages
pacman::p_load(pacman, knitr, tidyverse, readxl)
```

```{r xaringanExtra-clipboard, echo=FALSE}
# these allow any code snippets to be copied to the clipboard so they 
# can be pasted easily
htmltools::tagList(
    xaringanExtra::use_clipboard(
        button_text = "<i class=\"fa fa-clipboard\"></i>",
        success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    ),
    rmarkdown::html_dependency_font_awesome()
)
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```


# Purpose and Agenda

What do you need to be able to work with a range of datasets? This week focuses on core data wrangling skills; it is the first of four weeks that equip you with a set of skills focused on the dplyr and tidyr packages that are part of the tidyverse. The context is accessing and preparing data on colleges in advance of building a better rating system.

## What we'll do in this presentation

- Discussion
- Reading Recap
- Key Concept: select, rename, filter, arrange
- Key Concept: Ratings and rankings
- Code-along and data
- What's next: Assignment(s) and Readings

---

# Discussion

.panelset[

.panel[.panel-name[Background]



]

.panel[.panel-name[Discussion Question]

-

]

]

???

This discussion can serve as an opening to discussing what ggplot2's strengths are relative to the above two tools.

---

# Reading Recap

Substantive reading(s):

> Shelton, C. C., Koehler, M. J., Greenhalgh, S. P., & Carpenter, J. P. (2022). Lifting the veil on Teach ersP ayTeachers. com: An in vestigation of educational marketplace offerings and downloads. Learning, Media and Technology, 47(2), 268-287.

> Rosenberg, J. M., Borchers, C., Burchfield, M. A., Anderson, D., Stegenga, S. M., & Fischer, C. (2022). Posts about students on Facebook: a data ethics p erspective. Educational researcher, 51(8), 547-550.

Technical reading(s):

> R for Data Science, Data Transformation: https://r4ds.hadley.nz/data-transform

???

As in the first week, it may be good to pull out and summarize highlights from students' discussion posts here.

---

.title[Key Concept: select, rename, filter, arrange

.panelset[

.panel[.panel-name[xxx]


]

.panel[.panel-name[yyyy]


]

.panel[.panel-name[zzz]

]
]

---

# Key Concept: Ratings and rankings

.panelset[

.panel[.panel-name[xxx]


]

.panel[.panel-name[yyyy]


]

.panel[.panel-name[zzz]

]
]

---

# Code-along

.panelset[

.panel[.panel-name[select]

- and rename!
- yyy
-zzz

```{r, eval = FALSE, echo = TRUE}

```

]

.panel[.panel-name[filter]

- yyy
-zzz

```{r, eval = FALSE, echo = TRUE}

```

]

.panel[.panel-name[arrange]

- yyy
-zzz

```{r, eval = FALSE, echo = TRUE}

```

]

.panel[.panel-name[EDA]

- yyy
-zzz

```{r, eval = FALSE, echo = TRUE}

```

]

]

---

# What's next?

.panelset[


.panel[.panel-name[Weekly Assignment]

- 
- 

]

.panel[.panel-name[Data]

- We'll be using data from the educationdata package, specifically

]

.panel[.panel-name[Substantive Reading(s)]

- Substantive reading(s) for the week ahead:

> Park, J. J., Kim, B. H., Wong, N., Zheng, J., Breen, S., Lo, P., … & Poon, O. (2023). Inequality Beyond S tandardized Tests: Trends in Extr acurricular Activity Reporting in College A pplications Across Race and Class. Technical Report. Annenberg Institute at Brown University. <https ://edworkin g papers.>co m/ai23-749.

Linked in Canvas.

]

.panel[.panel-name[Technical Reading(s)]

- Technical reading(s) for the week ahead:

> Data Science in Education Using R, Using School-Level Aggregate Data to Illuminate Educational Inequities: https://datascienceineducation.com/c09

> R For Data Science, Strings: https://r4ds.had.co.nz/strings.html

Also linked in Canvas.

]

]